\section{Results}
\label{sec:results}

\subsection{Overview}
Briefly summarize the key experimental findings and highlight the main trends across datasets and implant schemes.

\subsection{Low-Level Fidelity Metrics}
Present quantitative results for SCAPE and all baselines.
\begin{itemize}
    \item Tables or bar plots comparing methods for each dataset.
    \item Highlight where SCAPE achieves the largest gains.
    \item Note any cases where baselines perform similarly or better.
\end{itemize}

\subsection{Representational Similarity Analysis}
Report RSA correlations between encoded and original image spaces.
\begin{itemize}
    \item Show RSA scatter plots or correlation heatmaps.
    \item Compare correlations across implant schemes and datasets.
    \item Discuss preservation of relational structure.
\end{itemize}

\subsection{Reconstruction Performance}
Present decoder-based reconstruction results.
\begin{itemize}
    \item Quantitative metrics for each dataset and encoding method.
    \item Example reconstructions for qualitative comparison.
    \item Highlight differences in how well structural detail is recovered.
\end{itemize}

\subsection{Ablation and Sensitivity Analyses (Optional)}
If performed, include:
\begin{itemize}
    \item Effect of removing SCAPE components (e.g., density adaptation, amplitude normalization).
    \item Sensitivity to key parameters (e.g., $\lambda$ in DoG).
\end{itemize}

\subsection{Summary of Results}
Provide a concise bullet-point recap of the main performance patterns and insights revealed by the experiments.
